\chapter{Testing Infrastructure}\label{chapter:infra}
All tests were performed in the \ac{AWS} us-east-2 Ohio region. The EC2 instances ran Ubuntu Server 
24.04 LTS Linux. Throughout the experiments, all the VMs were provisioned on dedicated hosts that 
are in the same availability zone and Virtual Private Cloud. This is particularly important 
for network I/O experiments, as the traffic between VMs sharing the same AZ and VPC is free of 
charge. We ran parallel benchmarks on general-purpose and compute-intensive dedicated hosts from 
the 5th, 6th, and 7th generations. We used dedicated hosts to ensure all nodes share 
the same physical server. This provides a controlled environment that allows the measurement 
of degradation caused by resource contention, without any external interference.\\
To deploy the resources for the different experiments, we used Terraform, which is an \ac{IaC} tool 
developed by HashiCorp. It allows defining and provisioning resources using the \ac{HCL}, 
ensuring the automation and reproducibility of the benchmarks. 
Additionally, we used Distexprunner \cite{distex}, which is a tool written in Python 
that facilitates writing and executing bash commands remotely across multiple nodes via
their public IPs. Our experiments generated JSON or CSV files that were gathered in an S3 
bucket using Distexprunner. We also implemented multiple scripts in Python using mainly the re 
package \cite{re} for parsing raw data and csv \cite{csv} for working with comma-separated values.
These scripts were essential to transform the raw data generated by our experiments into clean data 
that can be used for visualization and analysis. \\
We prepared different Amazon Machine Images (AMI) that have the  required 
software pre-installed (Python, Distexprunner, benchmarking tools, and monitoring utilities).
This minimizes the boot time of our instances and prevents configuration drift between runs. \\
These conditions apply to both the network and CPU experiments. The specific experimental 
setups will be discussed in detail within the methodology of each main section.